Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-21 vesicle_code/docs/adhesion/revisions/reviewer2.tex
finished reviewer 2 reubuttal

\begin{document}
\noindent
We thank the referee for the comments.  Below are our point-to-point responses to 
all the comments from the reviewer.  We have made revisions based on these comments.
We also note that a few typos from the original manuscript have
been corrected.  A separate list of changes can be found along with this list of responses.
\\ \\ \noindent 
Sincerely, \\ \noindent
Yuan-Nan Young, Shravan Veerapaneni, and Bryan Quaife

@@

\item\comment{Intro, 3rd paragraph: the buckling instability of aggregating RBC
membranes was not only predicted theroretically and numerically in
Flormann et al. [32] but also experimentally. In addition, a more
recent paper by Hoore et al. (Soft Matter 14, 6278-6289 (2018)) showed
the influence of the membrane shear elasticity on doublet shapes by 3D
modelling. This reference may be added.}
\begin{itemize}
  \item The paper by Hoore et al.~has been added to the introduction.
\end{itemize}

@@

\item\comment{In II. Governing equations and fig. 1: the right panel of the
figure and comments at the end of the section say that the part of
vesicle membranes that is out of the contact zone, including the parts
that are facing the outer fluid, experience an attractive interaction.
This is an artifact due to the long-range potential chosen by the
authors to model the interaction. In most experimental systems, the
adhesive forces leading to cluster formation are either due to direct
bridging of membranes or depletion phenomena that lead to osmotic
pressure gradients. This is the case for RBC aggregation for instance.
In both mechanisms, there is indeed no attractive force between membrane
parts that are out of the contact zone. It is perfectly fine to model
the interaction by an adhesion potential such as represented in fig 1
left, but its use introduces this artifact that should be acknowledged
by the authors. It is also due to the fact that the value of parameter
$\delta$ (the equilibrium distance) is not small compared to vesicle
dimensions. This is a numerical necessity to avoid the need to have
extremely refined meshes but again in experiments, the equilibrium
distance is usually much smaller that vesicle sizes and is more
comparable to molecular sizes (size of bridging molecules or of
depletion-inducing macromolecules). This deserves to be acknowledged
also, even though it may not impact too much the results.}
\begin{itemize}
  \item The referee is correct that the long-range attraction in the L.-J. is an artifact.
   In this work we focus on the adhesion effects on the hydrodynamics 
   of two vesicles drawn together by an external flow.
   Thus the L.-J. type potential is to keep the equilibrium distance, 
   following the works on red blood cells by Flormann {\em et al.}~Scientific
    Reports (2017) and Hoore {\em et al.}~Soft Matter (2018).  In both
    these papers, the Lennard-Jones potential is used to model the
    adhesive interaction between red blood cells.  
    Results from these modeling works are found to be in good agreement 
    with experiments and other models for adhesion (Ziherl and Svetina, PNAS (2017)).

@@

\item\comment{I cannot figure out why the external part of vesicles on profile 1
in fig 3-left is yellow. Shouldn't the force be weakly attractive in
that region, as in fig 1-right?}
\begin{itemize}
  \item In Figure 3 left, the color is not the adhesive force---it is the
    tension.  This is mentioned in the caption.  However, in the plot on
    the right that shows a zoomed in view of the contact zone, the
    adhesion force is plotted using the same method as in Figure 1.  In
    this case, the color is yellow only if the vesicle is sufficiently
    close to the other.
\end{itemize}

\item\comment{Figs 4-5 and related text: it would have been interesting to
compare results with the bending energy of vesicles without adhesion
potential in their equilibrium shape, to show the energy gain due to
adhesion (bending and total energy). This could be a way to compare
the total energy of the doublet to the energy of dissociated vesicles
and infer about the stability of doublets.}
\begin{itemize}
  \item We thank the referee for this suggestion. We did follow the referee's suggestion to check out this comparison. As pointed out by the referee in the seventh comment,
  the long-range attraction of the L-J potential that is used for membrane-membrane adhesion is not realistic and consequently the doublet always has lower total energy than
  dissociated vesicles. We are working on a more realistic adhesion potential for RBCs and we believe that this comparison will reveal more information there.
\end{itemize}

@@

\item\comment{ Figs 9(right) and 12(right): the stability diagrams are presented
in different parameter spaces for the extensional and shear flow
cases. It would have been interesting to see e.g. the diagram in the
interaction energy vs. flow strength $\chi$ for both cases, for
comparison.}
\begin{itemize}
  \item In the revision for the extensional example we first focus on how an initial vertical displacement 
  may affect the vesicle dynamics in a planar extensional flow (beginning of the new Section III). 
  There we find that a vesicle doublet is not likely to form unless the two vesicles are aligned
  on the converging axis of the extensional flow. In the fluid trap simulations we find that
    the reduced area has a large effect
    on determining if the doublet is broken or not, and on the
    inclination angle if the doublet is not broken.  Therefore, we have
    elected to investigate the effect of the reduced area for this
    example.

  \item For the shear flow simulations, we place the two vesicles on the same horizontal axis at $t=0$ (initial displacement is zero).
  Such head-on collision has been well studied for drops (clean or surfactant-laden), capsules, vesicles and red blood cells (without the L-J potential between membranes),
  and will serve as a good reference point to draw comparison with our results. For head-on collision between two vesicles,
  we find that  a smaller reduced area will increase the
    size of the contact area, and this will affect the critical
    Hamaker constant required to form a doublet. Thus we expect that the critical Hamaker constant for vesicle doublet formation reported will depend on the
    reduced area, vertical displacement and the shear rate. 
    To avoid
    diverting the focus of the paper, we have elected to divert an
    investigation of this behavior to future work.
\end{itemize}

@@

  \item There is a transient tank-treading motion that eventually (asymptotically) stops towards equilibrium. 
  Since we removed Figure 10 from the original version, we have chosen to
    not focus on the transient trank-treading motion and the vorticity and torque of the vesicles.
\end{itemize}

\item\comment{Fig 12: the authors did not comment the evolution of the
frequency of distance oscillations in doublets when the Hamaker constant
is varied. Can it be compared to the natural rotation frequency of the
shear flow $\chi/2$? Does a strongly adhesive doublet tend to behave as
a tumbling rigid object following something similar to Jeffery orbits?}
\begin{itemize}
  \item We now report the period of the doublets at three different
    shear rates and at a variety of Hamaker constants.

  \item For each of the shear rates and two of the Hamaker constants
    that we consider, we compute an effective aspect ratio which is the
    aspect ratio of a rigid body ellipse that would undergo the same
    period in the same shear flow.  The results indicate that the
    effective aspect ratio is independent of the shear rate, and this
    suggests that the dynamics of the doublet has similar properties to
    a rigid body undergoing a Jeffery orbit. 
\end{itemize}

\item\comment{Fig 12-right: is the stability limit between the red and blue
zones linear? This would mean that a stability criterion could take
the form of a dimensionless number involving the ratio of Hamaker
constant/shear rate, a potentially interesting result.}
\begin{itemize}
  \item We thank the referee for this observation.  We increased the
    parameter space of the phase diagram, and it does appear that there
    is a linear relationship.
  \item We have superimposed a straight line to show the transition
    between the two doublet and unbounded configurations.
  \item We also have repeated the phase diagram, but with a smaller
    separation distance.  With this new separation distance, the
    transition between the two regions appears to also be linear.
\end{itemize}

\item\comment{Fig 13 and comments about rheology: I am not sure to understand
the meaning of the left figure where it is shown that the intrinsic
viscosity of the doublet in suspension is about twice the viscosity of a
single vesicle. By definition, the intrinsic viscosity does not depend
on volume fraction, so I do not expect such a big difference between
bound vesicles and isolated vesicles. With the values chosen for the
Hamaker constant and the shear rate, it seems (from fig 12-left) that
the interaction is quite strong and that the distance between the two
vesicles in the doublet is always very short (close to the equilibrium
distance $\delta$). So the doublet must be roughly tumbling as a single
object and unless a co-rotating tank-treading motion of the two
membranes produces a strong dissipation by shear flow in the gap between
the two vesicles, the difference should not be so large. It is also
quite curious that the intrinsic viscosity of the doublet is so close to
exactly twice the viscosity of a single vesicle. Could the authors check
and elaborate on this ? Was the volume fraction properly scaled out
here? Also in the right panel, the bending/tension part is compared to
twice the intrinsic viscosity of a suspension of isolated vesicles. Why
twice ? Finally, it is not clear how technically the bending/tension and
adhesion parts are separated.  Overall I think the study on rheology
needs much more development than what is presented here with only one
set of parameter values and very little quantitative comments.}
\begin{itemize}
  \item The referee is correct that we miscalculated the intrinsic
    viscosity.  We had initially divided by only the area inside a
    single vesicle rather than both vesicles in the doublet.  This has
    been corrected, and now there is only a small increase in the
    intrinsic viscosity when compared to a single tank treading vesicle.

>>>>

\begin{document}
\noindent
We thank the referee for the comments.  Below are our point-to-point
responses to all the comments from the reviewer.  We have made revisions
based on these comments.  We also note that a few typos from the
original manuscript have been corrected.  A separate list of changes can
be found along with this list of responses.

\medskip
\noindent 
Sincerely, \\ \noindent
Yuan-Nan Young, Shravan Veerapaneni, and Bryan Quaife

@@

\item\comment{Intro, 3rd paragraph: the buckling instability of aggregating RBC
membranes was not only predicted theroretically and numerically in
Flormann et al.~[32] but also experimentally. In addition, a more
recent paper by Hoore et al.~(Soft Matter 14, 6278-6289 (2018)) showed
the influence of the membrane shear elasticity on doublet shapes by 3D
modelling. This reference may be added.}
\begin{itemize}
  \item The paper by Hoore et al.~has been added to the introduction.
\end{itemize}

@@

\item\comment{In II. Governing equations and fig. 1: the right panel of the
figure and comments at the end of the section say that the part of
vesicle membranes that is out of the contact zone, including the parts
that are facing the outer fluid, experience an attractive interaction.
This is an artifact due to the long-range potential chosen by the
authors to model the interaction. In most experimental systems, the
adhesive forces leading to cluster formation are either due to direct
bridging of membranes or depletion phenomena that lead to osmotic
pressure gradients. This is the case for RBC aggregation for instance.
In both mechanisms, there is indeed no attractive force between membrane
parts that are out of the contact zone. It is perfectly fine to model
the interaction by an adhesion potential such as represented in fig 1
left, but its use introduces this artifact that should be acknowledged
by the authors. It is also due to the fact that the value of parameter
$\delta$ (the equilibrium distance) is not small compared to vesicle
dimensions. This is a numerical necessity to avoid the need to have
extremely refined meshes but again in experiments, the equilibrium
distance is usually much smaller that vesicle sizes and is more
comparable to molecular sizes (size of bridging molecules or of
depletion-inducing macromolecules). This deserves to be acknowledged
also, even though it may not impact too much the results.}
\begin{itemize}
  \item The referee is correct that the long-range attraction in the
    L.-J.~is an artifact.  In this work we focus on the adhesion effects
    on the hydrodynamics of two vesicles drawn together by an external
    flow.  Thus the L.-J.~type potential is to keep the equilibrium
    distance, following the works on red blood cells by Flormann {\em et
    al.}~Scientific Reports (2017) and Hoore {\em et al.}~Soft Matter
    (2018).  In both these papers, the L.-J.~potential is used to model
    the adhesive interaction between red blood cells.  Results from
    these modeling works are found to be in good agreement with
    experiments and other models for adhesion (Ziherl and Svetina, PNAS
    (2017)).

@@

\item\comment{I cannot figure out why the external part of vesicles on
  profile 1 in fig 3-left is yellow. Shouldn't the force be weakly
  attractive in that region, as in fig 1-right?}
\begin{itemize}
  \item In Figure 3 left, the color is not the adhesive force---it is the
    tension.  This is mentioned in the caption.  However, in the plot on
    the right that shows a zoomed in view of the contact zone, the
    adhesive force is plotted using the same method as in Figure 1.  In
    this case, the color is yellow only if the vesicle is sufficiently
    close to the other.
\end{itemize}

\item\comment{Figs 4-5 and related text: it would have been interesting
  to compare results with the bending energy of vesicles without
  adhesion potential in their equilibrium shape, to show the energy gain
  due to adhesion (bending and total energy). This could be a way to
  compare the total energy of the doublet to the energy of dissociated
  vesicles and infer about the stability of doublets.}
\begin{itemize}
  \item We thank the referee for this suggestion. We did follow the
    referee's suggestion to check out this comparison. As pointed out by
    the referee in the seventh comment, the long-range attraction of the
    L.-J.~potential that is used for membrane-membrane adhesion is not
    realistic and consequently the doublet always has lower total energy
    than dissociated vesicles. We are working on a more realistic
    adhesion potential for RBCs and we believe that this comparison will
    reveal more information.
\end{itemize}

@@

\item\comment{Figs 9(right) and 12(right): the stability diagrams are
  presented in different parameter spaces for the extensional and shear
  flow cases. It would have been interesting to see e.g. the diagram in
  the interaction energy vs.~flow strength $\chi$ for both cases, for
  comparison.}
\begin{itemize}
  \item In the revision for the extensional example we first focus on
    how an initial vertical displacement may affect the vesicle dynamics
    in a planar extensional flow (beginning of the new Section III).
    There we find that a vesicle doublet is not likely to form unless
    the two vesicles are aligned on the converging axis of the
    extensional flow. In the fluid trap simulations we find that the
    reduced area has a large effect on determining if the doublet is
    broken or not, and on the inclination angle if the doublet is not
    broken.  Therefore, we have elected to investigate the effect of the
    reduced area for this example.

  \item For the shear flow simulations, we place the two vesicles on the
    same horizontal axis at $t=0$ (initial displacement is zero).  Such
    head-on collisions have been well-studied for drops (clean or
    surfactant-laden), capsules, vesicles, and red blood cells (without
    the L.-J.~potential between membranes), and will serve as a good
    reference point to draw comparison with our results. For head-on
    collisions between two vesicles, we find that  a smaller reduced
    area will increase the size of the contact area, and this will
    affect the critical Hamaker constant required to form a doublet.
    Thus we expect that the critical Hamaker constant for vesicle
    doublet formation reported will depend on the reduced area, vertical
    displacement and the shear rate.  To avoid diverting the focus of
    the paper, we have elected to divert an investigation of this
    behavior to future work.
\end{itemize}

@@

  \item There is a transient tank-treading motion that eventually
    (asymptotically) stops towards equilibrium.  Since we removed Figure
    10 from the original version, we have chosen to not focus on the
    transient tank-treading motion and the vorticity and torque of the
    vesicles.
\end{itemize}

\item\comment{Fig 12: the authors did not comment the evolution of the
  frequency of distance oscillations in doublets when the Hamaker
  constant is varied. Can it be compared to the natural rotation
  frequency of the shear flow $\chi/2$? Does a strongly adhesive doublet
  tend to behave as a tumbling rigid object following something similar
  to Jeffery orbits?}
\begin{itemize}
  \item We now report the period of the doublets at three different
    shear rates and at a variety of Hamaker constants and separation
    distances.

  \item For each of the shear rates and two of the Hamaker constants
    that we consider, we compute an effective aspect ratio which is the
    aspect ratio of a rigid body ellipse that would undergo the same
    period in the same shear flow.  The results indicate that the
    effective aspect ratio depends on the shear rate, and this suggests
    that the dynamics of the doublet in a shear flow is more complicated
    than Jeffery orbits.
\end{itemize}

\item\comment{Fig 12-right: is the stability limit between the red and
  blue zones linear? This would mean that a stability criterion could
  take the form of a dimensionless number involving the ratio of Hamaker
  constant/shear rate, a potentially interesting result.}
\begin{itemize}
  \item We thank the referee for this observation.  We increased the
    parameter space of the phase diagram, and it does appear that there
    is a linear relationship.

  \item We have superimposed a straight line to show the transition
    between the two doublet and unbounded configurations.

  \item We also have repeated the phase diagram, but with a smaller
    separation distance.  With this new separation distance, the
    transition between the two regions appears to also be linear.
\end{itemize}

\item\comment{Fig 13 and comments about rheology: I am not sure to
  understand the meaning of the left figure where it is shown that the
  intrinsic viscosity of the doublet in suspension is about twice the
  viscosity of a single vesicle. By definition, the intrinsic viscosity
  does not depend on volume fraction, so I do not expect such a big
  difference between bound vesicles and isolated vesicles. With the
  values chosen for the Hamaker constant and the shear rate, it seems
  (from fig 12-left) that the interaction is quite strong and that the
  distance between the two vesicles in the doublet is always very short
  (close to the equilibrium distance $\delta$). So the doublet must be
  roughly tumbling as a single object and unless a co-rotating
  tank-treading motion of the two membranes produces a strong
  dissipation by shear flow in the gap between the two vesicles, the
  difference should not be so large. It is also quite curious that the
  intrinsic viscosity of the doublet is so close to exactly twice the
  viscosity of a single vesicle. Could the authors check and elaborate
  on this? Was the volume fraction properly scaled out here? Also in the
  right panel, the bending/tension part is compared to twice the
  intrinsic viscosity of a suspension of isolated vesicles. Why twice?
  Finally, it is not clear how technically the bending/tension and
  adhesion parts are separated.  Overall I think the study on rheology
  needs much more development than what is presented here with only one
  set of parameter values and very little quantitative comments.}
\begin{itemize}
  \item The referee is correct that we miscalculated the intrinsic
    viscosity.  We had initially divided by only the area inside a
    single vesicle rather than both vesicles in the doublet.  This has
    been corrected, and now there is only a small increase in the
    intrinsic viscosity when compared to a single tank treading vesicle.
